Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-27 paper.tex
Edits to section 3

In this section, we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal
form of \citet{siek2012interpretations} to motivate the design of
hypercoercions.  We omit sum types and product types in this section
because \citet{siek2012interpretations} did not discuss them. We
assume a basic familiarity with coercions and direct readers who are
not familiar with them to \citet{siek2012interpretations} and
\citet{siek2015blame}.

@@

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $I!$ takes a value from an injectable type
$I$ to type $\TOOdyn$. An injectable type is simply a type that can be
cast directly to (injection) and from (projection) $\TOOdyn$. The
definition of injectable types depends on the blame strategy. With \lazyD,
all pre-types are injectable. With \lazyUD, only
$\POOfun{\TOOdyn}{\TOOdyn}$ is injectable.
%
A projection $I?^l$ takes a value from type $\TOOdyn$ to type $I$, or
halts the program and blames $l$ if the value is of a different type.
The coercion $\iota$ is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\ncSeq{c_1}{c_2})$ applies $c_1$ and
then $c_2$ in sequence. The function coercion $\ncFun{c_1}{c_2}$
applies $c_1$ to the argument of a function and $c_2$ to the return
value.

@@

\subsection{\lazyD{} hypercoercions}
\label{sec:ld-hc}

@@

\figref{fig:HC-D} defines the functions $id$, $seq$, and $cast$ that
construct \lazyD{} hypercoercions, towards satisfying the Cast abstract data 
type defined in Section~\ref{sec:framework}.
\figref{fig:HC-D} also defines some auxiliary functions.

@@

\subsection{\lazyUD{} hypercoercions}
\label{sec:lud-hc}

@@

The $cast$ function is defined in Figure~\ref{fig:HC-UD-cast}.  Prior
presentations of this function do not use auxiliary functions, as we
do here. The reason that we introduction the auxiliary functions
$\mathit{castToDyn}$, $\mathit{castFromDyn}$, $\mathit{castToInj}$,
and $\mathit{castFromInj}$, is to ensure that all of the are
structurally recursive, which makes them straightforward to define in
Agda.

@@

\item If the middle is of base types, then the remaining 62 bits are
  used to represent the head (25 bits), middle (11 bits), and tail (26
  bits).
  \begin{itemize}
  \item The head requires 1 bit to differentiate between $\epsilon$
    and $?^l$ and then 24 bits could be used for the label $l$.
  \item The middle would use 11 bits to differentiate all the base types.
  \item The tail requires 2 bits to differentiate between $\epsilon$,
    $!$, and $\bot^l$, and could use 24 bits for the label $l$.
  \end{itemize}
  In the event that a program requires too many blame labels, then one
  can fallback to using the heap-allocated structure for more
  coercions.
\end{itemize}

@@

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)=c$] constructs an identity cast from a type
  \item[$seq(c,c)=c$] composes two casts
  \item[$cast(T,l,T)=c$] constructs a cast from a source type, a label, and a 
  target type
  \item[$applyCast(v,c)=r$] applies a cast to a value, producing a cast result
  \end{description}
  We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

>>>>

In this section, we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal form of
\citet{siek2012interpretations} to motivate the design of hypercoercions.  We
omit sum types and product types in this section because
\citet{siek2012interpretations} did not discuss them. We assume a basic
familiarity with coercions, and suggest that readers unfamiliar with coercions
to familiarize themselves with \citet{siek2012interpretations} and
\citet{siek2015blame}.

@@

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $I!$ takes a value from an injectable type
$I$ to type $\TOOdyn$. An injectable type is simply a type that can be
cast directly to and from $\TOOdyn$. The
definition of injectable types depends on the blame strategy. With \lazyD,
all pre-types are injectable. With \lazyUD, only
$\POOfun{\TOOdyn}{\TOOdyn}$ and $\POOunit$ are injectable.
%
A projection $I?^l$ takes a value from type $\TOOdyn$ to type $I$, or
halts the program and blames $l$ if the value is of a different type.
The coercion $\iota$ is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\ncSeq{c_1}{c_2})$ applies $c_1$ and
then $c_2$ in sequence. The function coercion $\ncFun{c_1}{c_2}$
applies $c_1$ to the argument of a function and $c_2$ to the return
value.

@@

\subsection{\lazyD{} Hypercoercions}
\label{sec:ld-hc}

@@

\figref{fig:HC-D} defines the functions $id$, $seq$, and $cast$ that construct
\lazyD{} hypercoercions. We use these functions to satisfy the Cast abstract
data type defined in Section~\ref{sec:framework}. \figref{fig:HC-D} also defines
some auxiliary functions.

@@

\subsection{\lazyUD{} Hypercoercions}
\label{sec:lud-hc}

@@

The $cast$ function is defined in Figure~\ref{fig:HC-UD-cast}.  Prior
presentations of this function do not use auxiliary functions, as we
do here. The reason that we introduction the auxiliary functions
$\mathit{castToDyn}$, $\mathit{castFromDyn}$, $\mathit{castToInj}$,
and $\mathit{castFromInj}$, is to ensure that each of them is
structurally recursive, which makes them straightforward to define in
Agda.

@@

\item If the middle is of base types, then the remaining 62 bits are
  used to represent the head (25 bits), middle (11 bits), and tail (26
  bits).
  \begin{itemize}
  \item The head requires 1 bit to differentiate between $\epsilon$
    and $?^l$ and then 24 bits could be used for the label $l$.
  \item The middle would use 11 bits to differentiate all the base types.
  \item The tail requires 2 bits to differentiate between $\epsilon$,
    $!$, and $\bot^l$, and could use 24 bits for the label $l$.
  \end{itemize}
  The bits for blame labels represent an index into a table of blame information.
  In the event that a program requires too many blame labels, then the implementation
  can fallback to using the heap-allocated structure for more
  coercions.
\end{itemize}

@@

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)=c$] constructs an identity cast from a type
  \item[$seq(c,c)=c$] composes two casts
  \item[$cast(T,l,T)=c$] constructs a cast from a source type, a label, and a 
  target type
  \item[$applyCast(v,c)=r$] applies a cast to a value, producing a cast result
  \end{description}
  We use the syntax $c : T_1 \Longrightarrow T_2$ to mean $c$ is in the set $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}
